---
title: "Workload Model"
subtitle: "A summary of the workload model created for this organization"
title-block-banner: true
title-block-style: plain
format:
  html:
    toc: true
    toc-location: left
    toc-title: Contents
    css: styles.css
    code-fold: true
    output-file: 'index'
---

## Introduction

This report provides a summary view of data related to workers, tasks, and hours.

## Setup

### Environment



```{r setup}
#| message: false
here::i_am("final_report.qmd")
library(here)
library(ggplot2)
library(dplyr)
library(reactable)
library(reactablefmtr)
library(tippy)


cfg <- config::get()

ggplot2::theme_set(theme_bw())

mbcol <- "#1D5F97"

# TODO
# - add color pallette that aligns to malbay
# - add a label to work items graph to show what each mark means
# - figure out why rendering looks a little blurry
# - 90 degree angle of header row on heatmap or some other solution
# - Improve heatmap by
    # - making text smaller?
    # - Add tool tips with totals for columns and rows?
    # - 
# - After I get feedback on each view, start adding in data validation checks
# Can I adjust the scroll to scroll and not jump




# Random info: if prompted for gf pat, use credentials::set_github_pat()


```

### Reading in Data

The first data set is organized by Task and Position. 
It includes information about each of these combinations and a quick summary is 
provided below.

```{r}

# main_dat <- readr::read_csv(here(cfg$datapath, cfg$main), 
#                             show_col_types = FALSE)


main_dat <- readxl::read_xlsx(here(cfg$datapath, cfg$client_main)) %>% 
  setNames(make.names(names(main_dat)))
glimpse(main_dat)
```

The second data set organizes cases by tasks. 
It is meant to represent the full list of work items and the total number 
of hours that will be required to complete each.
This data will be compared to the first data set throughout.
An example of the output is printed below.

:::{.callout-note}

In our current example, this is from the same underlying data set.

:::



```{r}

# workitems_dat <- readr::read_csv(here(cfg$datapath, cfg$workitems),
#                                  show_col_types = FALSE)

workitems_dat <- main_dat %>% 
  select(Task, Total.Mins.Max, Department.Alignment) %>% 
  group_by(Task, Department.Alignment) %>% 
  summarize(Mins.Required = sum(Total.Mins.Max),
            instances = n())

glimpse(workitems_dat)

```

A final data set provides a similar set of information, but summed at the 
role level.

:::{.callout-note}

In our current example, this is from the same underlying data set.

:::

```{r}

# workitems_dat <- readr::read_csv(here(cfg$datapath, cfg$workitems),
#                                  show_col_types = FALSE)

role_dat <- main_dat %>% 
  select(Current.Position, Total.Mins.Max) %>% 
  group_by(Current.Position) %>% 
  summarize(Total.Mins = sum(Total.Mins.Max),
            Number.Tasks = n())

glimpse(role_dat)

```


## Preparing Data

This section outlines some data validation checks and a series of adjustments 
that are made to the input data based on global parameters. 

### Validating Data

:::{.callout-important}

TODO: I will work on this section as we confirm what the inputs will look like.
Again, I'm saving this for as final a version of the data we expect to collect
during this and future iterations of this work.

:::

Output from confirming data (probably a table)

- Number rows
- Total hours
- Total roles
- Total tasks
- Sum of planned hours should meet total number of fours after accounting for admin
- Sum of planned hours for each ind should meet expected supply, highlight above or below
- All individuals or roles should be included in the data set 
(placeholder inds can exist for planning)
- Sum of planned hours **within each period (week, month, year, etc.)** should
meet total number of hours after accounting for admin
- Sum of planned hours for **each individual or role within each period**
should meet total number of hours after accounting for administrative activities
- All tasks should be included, even if there is no alignment. 
The latter should be highlighted.
- Each ind's maximum and minimum dates should cover the full timeline. If they 
don't, the fact should be highlighted.

Other ideas?

### Adjusting Data

This section will outline any adjustments applied to the input data along with
a summary of any global parameters. 

A quick review of this data suggests a handful of adjustments. 
They include:

1. Drop items noted as "inaccurate".
2. Adjust to hours
3. Add period column to represent period

Once these adjustments are made, re-create the other two data sets.

```{r}

main_dat <-   main_dat %>% 
  filter(!stringr::str_detect(Task, 
                             stringr::regex("inaccurate", ignore_case = TRUE))) %>% 
  mutate(Total.Hours = round(Total.Mins.Max / 60, 2),
         period = "week")

workitems_dat <- main_dat %>% 
  select(Task, Total.Hours, Department.Alignment) %>% 
  group_by(Task, Department.Alignment) %>% 
  summarize(Hours.Required = sum(Total.Hours),
            Number.Tasks = n(),
            .groups = "drop")
role_dat <- main_dat %>% 
  select(Current.Position, Total.Hours) %>% 
  group_by(Current.Position) %>% 
  summarize(Total.Hours = sum(Total.Hours),
            Number.Tasks = n(), .groups = "drop")
```


:::{.callout-important}

OTHER IDEAS TO DISCUSS WITH TEAM:

1. Group tasks?

Some of the other generic ideas for this section includes:

- Overall timeline
- Standard work day/week/year
- Percent of worker's time that should be reserved for admin activities
- Total hours available after accounting for admin activities through the timeline
- Full list of workers/roles included in exercise
- Current quality level vs. optimal quality level

:::




## Views

### Hours by work item

The first view provides the hours by work item. 

```{r}

# prepare data
hours_by_item <- main_dat %>% 
  group_by(Task) %>% 
  summarize(Hours = sum(Total.Hours)) %>% 
  full_join(workitems_dat, by = join_by(Task)) %>% 
  mutate(Task = forcats::fct_infreq(Task, w = Hours.Required)) %>% 
  tidyr::replace_na(list(hours = 0))

xlab <- unique(main_dat$period)
stopifnot("period must be on the same time frame" = length(xlab) == 1)
  
  
# # Create graph
# hours_by_item %>% 
#   ggplot(aes(y = forcats::fct_rev(item))) +
#   geom_col(aes(x = hours),
#            fill = mbcol) +
#   geom_point(aes(x = hours_required)) +
#   labs(title = "Hours by work item",
#        y = "Work item", 
#        x = paste0("Hours per ", xlab))

part1 <- hours_by_item %>% 
  mutate(over_under_pct = Hours / Hours.Required-1) %>% 
  select(Task, Hours, Department.Alignment) %>% 
  ggplot(aes(y = forcats::fct_rev(Task))) +
  geom_col(aes(x = Hours, fill = Department.Alignment)) +
  labs(title = "Hours by task",
     y = "Task", 
     x = paste0("Hours per ", xlab)) + 
  scale_fill_manual(values = pals::kelly()) 
  # scale_fill_brewer(palette = "Set3")

  
part2 <- hours_by_item %>% 
  mutate(over_under_pct = Hours / Hours.Required-1) %>% 
  select(Task, over_under_pct, Department.Alignment) %>% 
  ggplot(aes(y = forcats::fct_rev(Task))) +
  geom_col(aes(x = over_under_pct, fill = Department.Alignment)) +
  scale_x_continuous(labels = scales::percent, limits = c(-1, 1),
                     breaks = c(-.75, 0, .75)) +
  labs(title = "% over/under",
     y = NULL, 
     x = "%")+
  theme(axis.text.y = element_blank()) + 
  scale_fill_manual(values = pals::kelly())
  # scale_fill_brewer(palette = "Set3")
  
library(patchwork)

part1 + part2 + plot_layout(widths = c(3, 1), guides = "collect") &
  theme(legend.position='bottom')

```


### Hours by individual and work item

This section shows the expect time for each task grouped by individuals or roles.

```{r}
#| fig-width: 13.5
#| fig-height: 6.6
#| column: screen-inset-right


# prepare data
hours_by_item_role <- main_dat %>%
  full_join(workitems_dat %>% select(Task, Hours.Required), by = join_by(Task)) %>% 
  filter(!is.na(Total.Hours)) %>% 
  group_by(Task, Current.Position, Department.Alignment) %>%
  summarize(hours = sum(Total.Hours), .groups = "drop") %>% 
  # full_join(workitems_dat) %>%
  mutate(item = forcats::fct_infreq(Task, w = hours)) #%>%
  # tidyr::replace_na(list(hours = 0))

xlab <- unique(main_dat$period)
stopifnot("period must be on the same time frame" = length(xlab) == 1)
  
  
# Create graph
hours_by_item_role %>% 
  ggplot() +
  geom_col(aes(x = hours, y = forcats::fct_rev(Task), fill = Department.Alignment), 
           position = "dodge") +
  facet_wrap(.~Current.Position, scale = "free_y", nrow = 3) +
  labs(title = "Hours by work item",
       y = "Work item", 
       x = paste0("Hours per ", xlab)) +
  theme(legend.position = "bottom") + 
  scale_fill_manual(values = pals::kelly())



```

### Gantt Chart

This view provides the time expected on each task throughout the time period.

This is not relevant for these data.

```{r}
#| fig-width: 14.6
#| fig-height: 7
#| column: screen-inset-right
# main_dat %>%
#   # filter(role == "manager") %>%
#   ggplot(aes(x=lubridate::ymd(start_dt), xend=lubridate::ymd(end_dt), y=item, yend=item)) +
#   # theme_bw()+ #use ggplot theme with black gridlines and white background
#   geom_segment(linewidth=3,
#                color = mbcol) + #increase line width of segments in the chart
#   facet_wrap(.~role, scales = "free_y") +
#   labs(title='Schedule', x='Time', y='Work Item')

```

### Heatmap

Finally, this view highlights where time is being spent by the team and 
where there are potential gaps.

```{r}
#| column: screen-inset-right

# https://glin.github.io/reactable/articles/examples.html#basic-usage

# TODO: 
# Add some data that takes into account higher level work tasks / skills
hm_dat <- main_dat %>% 
  select(item, exp_hours, role) %>% 
  tidyr::pivot_wider(names_from = role, 
                     values_from = exp_hours) %>% 
  as.data.frame()
 

with_tooltip <- function(value, tooltip, ...) {
  htmltools::div(style = "text-decoration: underline; text-decoration-style: dotted; cursor: help",
      tippy(value, tooltip, ...))
}


reactable(hm_dat, 
          resizable = TRUE,
          pagination = FALSE,
          bordered = TRUE,
          wrap = FALSE,
                    # width = 2000,
          # style = list(maxWidth = 3000),
          defaultColDef = colDef(footer = function(values) sum(values, na.rm = T),
                                 footerStyle = list(fontWeight = "bold"),
                                 minWidth = 100,
                                 style = color_scales(hm_dat, 
                                                      colors = RColorBrewer::brewer.pal(5, "Oranges"))),
          columns = list(
            item = colDef(minWidth = 200, filterable = TRUE, 
                         sticky = "left", footer = "Total"),
            # There must be a better way to do this...
            # cat(paste0("`", names(hm_dat), 
            #            "` = colDef(header = with_tooltip('",names(hm_dat),
            #            "','",names(hm_dat),"'))"), 
            #     sep = ",\n")
            # Maybe pass in a function to finish creating this table...?
            `manager` = colDef(header = with_tooltip('manager','manager')),
            `shift supervisor 1` = colDef(header = with_tooltip('shift supervisor 1','shift supervisor 1')),
            `shift supervisor 2` = colDef(header = with_tooltip('shift supervisor 2','shift supervisor 2')),
            `theater ops specialist sr 1` = colDef(header = with_tooltip('theater ops specialist sr 1','theater ops specialist sr 1')),
            `theater ops specialist sr 2` = colDef(header = with_tooltip('theater ops specialist sr 2','theater ops specialist sr 2')),
            `theater ops specialist sr 3` = colDef(header = with_tooltip('theater ops specialist sr 3','theater ops specialist sr 3')),
            `theater ops specialist 1` = colDef(header = with_tooltip('theater ops specialist 1','theater ops specialist 1')),
            `theater ops specialist 2` = colDef(header = with_tooltip('theater ops specialist 2','theater ops specialist 2')),
            `theater ops specialist 3` = colDef(header = with_tooltip('theater ops specialist 3','theater ops specialist 3'))
            )
          )
```


## Other ideas

- A view that highlights unfulfilled hours or work items

## Questions that need to be answered in report
- How many people would it take to do 100% of the required work?
- How do things change as the quality of work changes (increases or decreases)?


# Notes from discussion on 4/21/2023
- Want another column in to categorize
  - Skill sets within the Work Item table
  - Jerry asked about whether this is useful for aggregation or filtering?
-The ability to filter in column will be a benefit
- 

### Next Steps
- I will focus hard on the heatmap and finalizing the view
- Reduce the font size on heat map
- add in grouped skill for each item
- Add ability to switch between number or name or title

# Future work...
- Add in ability to organize graphs by teams


